### 2023-10-03 outputs/paper/paper.qmd
Complete data cleaning

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Yufei Liu
thanks: "Code and data are available at: https://github.com/Florence-Liu/bicycle-theft"
date: "`r Sys.time()`"
date-format: " MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-10-03 scripts/02-cleaning_data.R

@@ -1,0 +1,69 @@
+#### Preamble ####
+# Purpose: Cleans the Bicycle Theft data from OpenDataToronto
+# Author: Yufei Liu
+# Date: 2 Oct 2023
+# Contact: florence.liu@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+
+#### Read in the raw data ####
+raw_data <- read_csv("inputs/data/bicycle_theft.csv", show_col_types = FALSE)
+
+
+#### Clean data ####\
+cleaned_data <-
+  raw_data |>
+  janitor::clean_names() |> # to make variables name easier to read
+  # select our variable of interest
+  select(occ_year, occ_date, report_date, premises_type, status) |> 
+  tidyr::drop_na()
+
+cleaned_data <- 
+  # add a new column showing the difference in days between the report date and occurrence date
+  # add a new column showing the numerical expression of occurrence month as a character
+  cleaned_data |> mutate(difference = as.numeric(report_date - occ_date),
+                         occ_month = as.character(month(occ_date))) |>
+  # filter to get thefts of which occurrence year between 2014 and 2022
+  filter(occ_year >= 2014 & occ_year <= 2022) |>
+  # make the numerical occurrence yeaer as a character
+  mutate(occ_year = as.character(occ_year)) |>
+  # rename columns to be more readable
+  rename(location = premises_type, bicycle_status = status) |>
+  # drop columns that will not be used in the analysis
+  select(occ_year, occ_month, difference, location, bicycle_status)
+
+
+#### Data validation ####
+# Referencing data from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+
+
+# Check that occurrence_year are all between 2014 and 2022
+cleaned_data$occ_year |> as.numeric() |> max() <= 2022
+cleaned_data$occ_year |> as.numeric() |> min() >= 2014
+
+
+# Check that occurrence_month are all between 1 and 12
+cleaned_data$occ_month |> as.numeric() |> max() <= 12
+cleaned_data$occ_month |> as.numeric() |> min() >= 1
+
+
+# Check that the difference in days between the occurrence date and report date is positive
+cleaned_data$difference |> min() <= 0
+
+
+
+# Check variable types
+cleaned_data$occ_year |> class() == "character"
+cleaned_data$occ_month |> class() == "character"
[…]

### 2023-10-03 scripts/02-cleaning_data.R
Correct typo in comments

@@ -29,7 +29,7 @@ cleaned_data <-
                          occ_month = as.character(month(occ_date))) |>
   # filter to get thefts of which occurrence year between 2014 and 2022
   filter(occ_year >= 2014 & occ_year <= 2022) |>
-  # make the numerical occurrence yeaer as a character
+  # make the numerical occurrence year as a character
   mutate(occ_year = as.character(occ_year)) |>
   # rename columns to be more readable
   rename(location = premises_type, bicycle_status = status) |>

### 2023-10-03 outputs/paper/paper.qmd
Initial data presentation

@@
 
-## Quarto
+```{r}
+#| echo: false
+#| message: false
+#| warning: false
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+#### Workspace setup ####
+library(tidyverse)
+library(knitr)
 
-## Running Code
+#### Read in the cleaned data ####
+data <- read_csv(here::here("outputs/data/cleaned_bicycle_theft.csv"), show_col_types = FALSE)
+```
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+# Introduction
+
+You can and should cross-reference sections and sub-sections.
+
+The remainder of this paper is structured as follows. @sec-data....
+
+# Data {#sec-data}
 
 ```{r}
-1 + 1
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
+
+data |> group_by(occ_year) |> summarise(total = n()) |> kable()
 ```
 
-You can add options to executable code like this
+Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
 ```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
 #| echo: false
-2 * 2
+
+data |> ggplot(aes(x=occ_year, fill=location)) +
+  geom_bar()
+
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+Talk more about it.
+
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+
+```{r}
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> ggplot(aes(x=occ_year, fill=bicycle_status)) +
+  geom_bar()
+```
+
+```{r}
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> filter(difference <= 100) |> 
+  ggplot(aes(x=occ_year, y=difference, color=occ_month)) +
+  geom_point()
+```
+```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
+
+data |> group_by(occ_month) |> summarise(total = n()) |> kable()
+
+```
+
+\newpage
+
+# References

### 2023-10-03 scripts/02-cleaning_data.R
Change test for variable difference

@@ -51,8 +51,8 @@ cleaned_data$occ_month |> as.numeric() |> max() <= 12
 cleaned_data$occ_month |> as.numeric() |> min() >= 1
 
 
-# Check that the difference in days between the occurrence date and report date is positive
-cleaned_data$difference |> min() <= 0
+# Check that the difference in days between the occurrence date and report date is non-negative
+cleaned_data$difference |> min() >= 0
 
 
 

### 2023-10-03 scripts/02-cleaning_data.R
Add requirement for variable difference

@@ -29,6 +29,8 @@ cleaned_data <-
                          occ_month = as.character(month(occ_date))) |>
   # filter to get thefts of which occurrence year between 2014 and 2022
   filter(occ_year >= 2014 & occ_year <= 2022) |>
+  # filter to get reported date within one year of occurrence date
+  filter(difference <= 365) |>
   # make the numerical occurrence year as a character
   mutate(occ_year = as.character(occ_year)) |>
   # rename columns to be more readable
@@ -51,8 +53,9 @@ cleaned_data$occ_month |> as.numeric() |> max() <= 12
 cleaned_data$occ_month |> as.numeric() |> min() >= 1
 
 
-# Check that the difference in days between the occurrence date and report date is non-negative
+# Check that the difference in days between the occurrence date and report date are all between 0 and 365
 cleaned_data$difference |> min() >= 0
+cleaned_data$difference |> max() <= 365
 
 
 
@@ -67,3 +70,4 @@ cleaned_data$bicycle_status |> class() == "character"
 
 #### Save data ####
 write_csv(cleaned_data, "outputs/data/cleaned_bicycle_theft.csv")
+

### 2023-10-04 scripts/02-cleaning_data.R
Change variable occ_month

@@ -19,14 +19,13 @@ cleaned_data <-
   raw_data |>
   janitor::clean_names() |> # to make variables name easier to read
   # select our variable of interest
-  select(occ_year, occ_date, report_date, premises_type, status) |> 
+  select(occ_year, occ_date, occ_month, report_date, premises_type, status) |> 
   tidyr::drop_na()
 
 cleaned_data <- 
   # add a new column showing the difference in days between the report date and occurrence date
   # add a new column showing the numerical expression of occurrence month as a character
-  cleaned_data |> mutate(difference = as.numeric(report_date - occ_date),
-                         occ_month = as.character(month(occ_date))) |>
+  cleaned_data |> mutate(difference = as.numeric(report_date - occ_date)) |>
   # filter to get thefts of which occurrence year between 2014 and 2022
   filter(occ_year >= 2014 & occ_year <= 2022) |>
   # filter to get reported date within one year of occurrence date
@@ -49,8 +48,9 @@ cleaned_data$occ_year |> as.numeric() |> min() >= 2014
 
 
 # Check that occurrence_month are all between 1 and 12
-cleaned_data$occ_month |> as.numeric() |> max() <= 12
-cleaned_data$occ_month |> as.numeric() |> min() >= 1
+cleaned_data$occ_month |> unique() |> sort() == c("April","August","December","February",
+                                                 "January","July","June","March",    
+                                                 "May","November","October","September")
 
 
 # Check that the difference in days between the occurrence date and report date are all between 0 and 365

### 2023-10-04 outputs/paper/paper.qmd
Complete paper

--- 2023-10-03 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
-author: 
-  - Yufei Liu
+title: "Public security, law enforcement, and bicycle theft in Toronto"
+author: "Yufei Liu"
 thanks: "Code and data are available at: https://github.com/Florence-Liu/bicycle-theft"
@@
 date-format: " MMMM YYYY"
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Bicycle theft has been a perennial issue for city safety in Toronto.
+The analysis is conducted based on data on opendatatoronto including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service.
+There exists a decreasing trend and low recovery rates in overall offence in recent years even people tend to report immediately, a seasonal pattern that more offence occurred in summer, and a geographical pattern that outdoor theft decreased.
+It implies possible ineffectiveness of tracking and security management, and further work could look at influencing factors contributing to the patterns."
 format: pdf
@@
 
-## Quarto
+# Introduction
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+Cycling is one of the most popular transportation methods and relaxing activities in Toronto and contributes to reducing traffic congestion.
+The City of Toronto is working to support cycling by building infrastructures, such as bike lane and bike parking facilities, and create a cycling-friendly environment [@citeThecity].
+However, together with the prevalence of cycling, bicycle theft became a major concern for cyclists and a perennial problem for the city safety in Toronto.
 
-## Running Code
+In this paper, we use R [@citeR] to analyze `bicycle theft` dataset [@citeData] in `opendatatoronto` [@citeOpendatatoronto] to discover any potential time and geographical patterns in bicycle theft occurrence in Toronto from 2014 to 2022.
+We explore time and location variables as well as recovery rate and report habit through bar plots, scatter plots, and tables.
+We find that the overall trend of bicycle theft in Toronto decreased in recent year with a potential seasonal pattern that more thefts occurred in summer, and a geographical pattern that outdoor theft decreased, along with relatively law recovery rate and immediate report habit.
+Biases exist in the dataset and may influence some analysis.
+Further work could explore possible reasons behind the pattern combining with other dataset.
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
+The remainder of this paper is structured as follows: @sec-data discusses the data with **1** @sec-data_introduction includes basic information about data collection, and discussion about biases and ethical concern about the data, @sec-data_cleaning includes packages used to clean the data and cleaning process with information about variables of interest, and @sec-data_analysis includes graphs and tables representing relationships between variable, and some discussion about implications.
 
-```{r}
-1 + 1
-```
+# Data {#sec-data}
 
-You can add options to executable code like this
+## Data introduction  {#sec-data_introduction}
+
+The dataset used in this analysis was obtained from Toronto's Open Data Portal and can be accessed from the package `opendatatoronto` [@citeOpendatatoronto].
+The dataset `Bicycle Theft` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022, including detailed information about the theft such as location, bike basic information, etc [@citeData].
+The dataset has been modified several times to update the data, and in this analysis, we will use the latest `csv` version refreshed in Apr 4, 2023.
+
+Due to privacy consideration, Toronto Police Service has artificially blurred the detailed location of occurrence in this data to protect parties involved and locations were offset to the nearest road intersection [@citeData].
+In this case, the accuracy of location may be impacted and become less imformative.
+
+Since the occurrence of bicyle theft was self-reported, the reporter may choose to hide or inaccurately report some information such as detailed information about the bike due to privacy consideration as lost properties are not actively investigated by the Toronto Police Service and reports are only used for records purposes[@citeTorontopoliceservice].
+It is also possible that people forget the exact time of the occurrence and detailed information about the bike if things happened long time ago.
+All these factors contribute to the existence of reporting bias, and make our data biased and less accurate for some analysis.
+
+## Data cleaning {#sec-data_cleaning}
+
+We will use `R` to do the analysis in this report [@citeR].
+After acquiring the dataset from package `opendatatoronto`[@citeOpendatatoronto], we used package `tidyverse`[@citeTidyverse] and package `janitor`[@citeJanitor] to clean the data and select variables for further analysis.
+
+We selected variables from the raw data [@citeData]:
+
+* occ_year: Year offence occurred
+
+* occ_month: Month offence occurred
+
+* occ_date: Date offence occurred
+
+* report_date: Date offence occurred
+
+* premises_type: Premises type of offence
+
+* status: Whether the bicycle is recovered, stolen, or unknown
+
+To analyze the relationship between occurrence date and report date of offence, we constructed a new numerical variable `difference` to represent the difference in days between the report date and occurrence date, and filter any offence whose difference is larger than 365, that means we will not analyze offence that is not reported within one year from the occurrence date.
+The number of offence that reports long time later is quite small and may not be representative due to possible inaccurate report information.
+
+After changing the type and name of some variables, filtering offence occurred between 2014 and 2022, and removing missing values, we selected our final variable for analysis:
+
+* occ_year: Year offence occurred
+
+* occ_month: Month offence occurred
+
+* difference: Difference in days between occurrence date and report date
+
+* location: Premises type of offence
+
+* bicycle_status: Whether the bicycle is recovered, stolen, or unknown
+
+## Data analysis {#sec-data_analysis}
+
+In the analysis, we will use package `knitr` [@citeKnitr], package `here` [@citeHere], and package `formattable` [@citeFormattable] to load the cleaned data and make figures and tables to explore relationship between variables.
 
@@
 #| echo: false
-2 * 2
+#| message: false
+#| warning: false
+
+#### Workspace setup ####
+library(tidyverse)
+library(knitr)
+library(formattable)
+
+#### Read in the cleaned data ####
+data <- read_csv(here::here("outputs/data/cleaned_bicycle_theft.csv"), 
+                 show_col_types = FALSE,
+                 col_types = cols(occ_year = col_character(),
+                                  occ_month = col_character()))
 ```
 
-The `echo: false` option disables the printing of code (only output is displayed).
+@fig-yearly_count_by_month shows monthly number of offence for each year from 2014 to 2022.
+The bar plot provides a clear demonstration for not only the overall trend of bicycle theft through years, but also the potential seasonal pattern that could be further explored.
+It shows that the total number of offence in a year increased before 2018 and began decrease since 2018.
+The change is relatively not significant within 1000, and the number of offence in 2022 is similar to that in 2014.
+Possible reasons include improvement of security equipment of the bike, the market behind bicycle theft influenced by pandemic, and people's usage habit of bike.
+These implications need more data to analyze and could be our further interest.
+
+@fig-yearly_count_by_month also illustrates potential seasonal pattern that there much less bicycle theft in winter, eg.
+December, January, and February.
+The offence increases a lot in other seasons especially in summer, eg.
+June, July, and August.
+This may be related to the weather in Toronto that many people will not ride bikes in winter due to extremely freezing weather and frequent snowfall.
+
+```{r}
+#| label: fig-yearly_count_by_month
+#| fig-cap: Bar plot of monthly number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> ggplot(aes(x=occ_year, fill=occ_month)) +
+  geom_bar() +
+  labs(x = "Occurrence Year", y = "Number of occurrence", fill = "Occurrence Month") +
+  scale_x_discrete() +
+  theme_classic()
+```
+
+Other than seasonal pattern, @fig-location_count_by_year also demonstrates the relationship between the location type and number of bicycle theft for each year.
+It clearly shows that bicycles were mostly stolen when parked outside, and this aligns with common sense that outdoor parking is more vulnerable to theft.
+However, it is kind of surprising that the second most common location for bicycle theft is apartment instead of house.
+Residence transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
+For residence in house, they may rather choose vehicles that are faster such as cars.
+
+@fig-location_count_by_year also illustrates that the number of offence happened outside decreases while thefts happened indoor especially in residence increases in recent years.
+Factors contribute to the phenomenon could be further explored combining other data.
+
+It is noticeable that there are many reports with location type `Other`.
+Since it is not specified and the amount of `Other` type location cannot be neglected, how we treat the data may influence the analysis, and there may exists possible reporting bias.
+
+```{r}
+#| label: fig-location_count_by_year
+#| fig-cap: Bar plot of number of occurrence of bicycle theft in difference type of location in Toronto for each year from 2014 to 2022.
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> ggplot(aes(x=location, fill=occ_year)) +
+  geom_bar() +
+  labs(x = "Location Type", y = "Number of occurrence", fill = "Occurrence Year") +
+  scale_x_discrete() +
+  theme_classic()
+
+```
+
+Other than external pattern, we are also interested how internal variable associated with bicycle theft.
+@fig-diff_count_by_year shows the monthly difference in days between occurrence date and report date for each year from 2014 to 2022.
+It illustrates that people tend to report the offence within the same day of occurrence since there are more scatters near 0.
+Combined with @fig-yearly_count_by_month, there are much fewer thefts happened in December, however, the interval between occurrence and report date is much shorter in December since there are much more dots representing December near 0.
+
+```{r}
+#| label: fig-diff_count_by_year
+#| fig-cap: Scatter plot of monthly difference in days between occurrence date and report date for bicycle thefts in Toronto for each year from 2014 to 2022.
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> 
+  ggplot(aes(x=occ_year, y=difference,
+             color=occ_month)) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Occurrence Year", y = "Difference in days between occurrence date and report date", fill = "Occurrence Month") +
+  theme_classic()
+```
+
+@tbl-rate_by_year and @tbl-rate_by_month contains recovery rate of bicycle in each year and month respectively.
+The recovery rate is calculated as the percentage of stolen bicycle that could be successfully returned to the owner.
+It does not have a clear trend of recovery rate through years, however, 2022 has the lowest recovery rate, which is much less than that in other years (See @tbl-rate_by_year).
+This is not a good signal that the three lowest recovery rates come from 2022, 2019, 2021, although the highest recovery rate comes from 2020.
+Such unusual pattern is worth investigating.
+
+@tbl-rate_by_month shows that the recovery rate in winter is lower than that in summer in general.
+Consider potential seasonal pattern for bicycle theft in a year, this may also be related to the weather in Toronto that severe and freezing weather in winter constrains people's activity in both positive and negative way.
+Even recovery rate is higher in summer, it is still not a high value that largest recovery rate for each year is 17.58% and for each month is 15.70%.
+
+```{r}
+#| label: tbl-rate_by_year
+#| tbl-cap: Yearly recovery rate of bicycle theft in Toronto from 2014 to 2022.
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> group_by(occ_year) |> 
+  summarise(rate = percent(sum(bicycle_status == "RECOVERED")/nrow(data)*100)) |> 
+  arrange(rate) |> 
+  kable(col.names = c("Year", "Recovery Rate"))
+
+```
+
+```{r}
+#| label: tbl-rate_by_month
+#| tbl-cap: Monthly recovery rate of bicycle theft in Toronto from 2014 to 2022
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> group_by(occ_month) |> 
+  summarise(rate = percent(sum(bicycle_status == "RECOVERED")/nrow(data)*100)) |> 
+  arrange(rate) |> 
+  kable(col.names = c("Month", "Recovery Rate"))
+
+```
+
+In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
+Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
+Bicycle theft is common in Toronto but it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
+
+\newpage
+
+# References

### 2023-10-04 outputs/paper/paper.qmd
Correct typos and adjust graph axis size

@@
 abstract: "Bicycle theft has been a perennial issue for city safety in Toronto.
-The analysis is conducted based on data on opendatatoronto including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service.
-There exists a decreasing trend and low recovery rates in overall offence in recent years even people tend to report immediately, a seasonal pattern that more offence occurred in summer, and a geographical pattern that outdoor theft decreased.
+The analysis is conducted based on data on opendatatoronto portal including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service.
+There exists a decreasing trend and low recovery rates in overall offences in recent years even people tend to report immediately, a seasonal pattern that more offences occurred in summer, and a geographical pattern that outdoor bicycle thefts decreased.
 It implies possible ineffectiveness of tracking and security management, and further work could look at influencing factors contributing to the patterns."
@@
 Cycling is one of the most popular transportation methods and relaxing activities in Toronto and contributes to reducing traffic congestion.
-The City of Toronto is working to support cycling by building infrastructures, such as bike lane and bike parking facilities, and create a cycling-friendly environment [@citeThecity].
+The City of Toronto is working to support cycling by building infrastructures, such as bike lanes and bike parking facilities, and create a cycling-friendly environment [@citeThecity].
 However, together with the prevalence of cycling, bicycle theft became a major concern for cyclists and a perennial problem for the city safety in Toronto.
 
-In this paper, we use R [@citeR] to analyze `bicycle theft` dataset [@citeData] in `opendatatoronto` [@citeOpendatatoronto] to discover any potential time and geographical patterns in bicycle theft occurrence in Toronto from 2014 to 2022.
+In this paper, we use R [@citeR] to analyze `bicycle thefts` dataset [@citeData] in `opendatatoronto` package [@citeOpendatatoronto] to discover any potential time and geographical patterns in bicycle theft occurrence in Toronto from 2014 to 2022.
 We explore time and location variables as well as recovery rate and report habit through bar plots, scatter plots, and tables.
-We find that the overall trend of bicycle theft in Toronto decreased in recent year with a potential seasonal pattern that more thefts occurred in summer, and a geographical pattern that outdoor theft decreased, along with relatively law recovery rate and immediate report habit.
+We find that the overall occurrence of bicycle theft in Toronto decreased in recent years with a potential seasonal pattern that more thefts occurred in summer, and a geographical pattern that outdoor thefts decreased, along with relatively law recovery rates and immediate report habit.
 Biases exist in the dataset and may influence some analysis.
-Further work could explore possible reasons behind the pattern combining with other dataset.
+Further work could explore possible reasons behind the patterns combining with other dataset.
 
-The remainder of this paper is structured as follows: @sec-data discusses the data with **1** @sec-data_introduction includes basic information about data collection, and discussion about biases and ethical concern about the data, @sec-data_cleaning includes packages used to clean the data and cleaning process with information about variables of interest, and @sec-data_analysis includes graphs and tables representing relationships between variable, and some discussion about implications.
+The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection, and discussion about biases and ethical concerns about the data, @sec-data_cleaning includes packages used to clean the data and cleaning process with information about variables of interest, and @sec-data_analysis includes graphs and tables representing relationships between variables, and some discussions about implications.
 
@@
 The dataset used in this analysis was obtained from Toronto's Open Data Portal and can be accessed from the package `opendatatoronto` [@citeOpendatatoronto].
-The dataset `Bicycle Theft` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022, including detailed information about the theft such as location, bike basic information, etc [@citeData].
-The dataset has been modified several times to update the data, and in this analysis, we will use the latest `csv` version refreshed in Apr 4, 2023.
+The dataset `Bicycle Thefts` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022, including detailed information about the offence such as location, bike basic information, etc [@citeData].
+The dataset has been modified several times to update the data, and in this analysis, we will use the latest `csv` version refreshed on Apr 4, 2023.
 
 Due to privacy consideration, Toronto Police Service has artificially blurred the detailed location of occurrence in this data to protect parties involved and locations were offset to the nearest road intersection [@citeData].
-In this case, the accuracy of location may be impacted and become less imformative.
+In this case, the accuracy of location may be impacted and become less informative.
 
-Since the occurrence of bicyle theft was self-reported, the reporter may choose to hide or inaccurately report some information such as detailed information about the bike due to privacy consideration as lost properties are not actively investigated by the Toronto Police Service and reports are only used for records purposes[@citeTorontopoliceservice].
+Since the occurrence of bicycle theft was self-reported, the reporter may choose to hide or inaccurately report some information such as detailed information about the bike due to privacy consideration as lost properties are not actively investigated by the Toronto Police Service and reports are only used for records purposes[@citeTorontopoliceservice].
 It is also possible that people forget the exact time of the occurrence and detailed information about the bike if things happened long time ago.
@@
 
-* report_date: Date offence occurred
+* report_date: Date offence reported
 
@@
 
-To analyze the relationship between occurrence date and report date of offence, we constructed a new numerical variable `difference` to represent the difference in days between the report date and occurrence date, and filter any offence whose difference is larger than 365, that means we will not analyze offence that is not reported within one year from the occurrence date.
+To analyze the relationship between occurrence date and report date of offence, we constructed a new numerical variable `difference` to represent the difference in days between the report date and occurrence date, and filter any offences whose difference is larger than 365, that means we will not analyze offence that is not reported within one year from the occurrence date.
 The number of offence that reports long time later is quite small and may not be representative due to possible inaccurate report information.
@@
 @fig-yearly_count_by_month shows monthly number of offence for each year from 2014 to 2022.
-The bar plot provides a clear demonstration for not only the overall trend of bicycle theft through years, but also the potential seasonal pattern that could be further explored.
+The bar plot provides a clear demonstration for not only the overall trend of bicycle thefts through years, but also the potential seasonal pattern that could be further explored.
 It shows that the total number of offence in a year increased before 2018 and began decrease since 2018.
@@
 
-@fig-yearly_count_by_month also illustrates potential seasonal pattern that there much less bicycle theft in winter, eg.
+@fig-yearly_count_by_month also illustrates potential seasonal pattern that there are much fewer bicycle thefts in winter, eg.
 December, January, and February.
@@
 It clearly shows that bicycles were mostly stolen when parked outside, and this aligns with common sense that outdoor parking is more vulnerable to theft.
-However, it is kind of surprising that the second most common location for bicycle theft is apartment instead of house.
-Residence transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
-For residence in house, they may rather choose vehicles that are faster such as cars.
+However, it is surprising that the second most common location for bicycle theft is apartment instead of house.
+Residents' transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
+For residents in house, they may rather choose vehicles that are faster such as cars to commute.
 
-@fig-location_count_by_year also illustrates that the number of offence happened outside decreases while thefts happened indoor especially in residence increases in recent years.
-Factors contribute to the phenomenon could be further explored combining other data.
+@fig-location_count_by_year also illustrates that the number of offence happened outside decreased while thefts happened indoor especially in residence increased in recent years.
+Factors contributing to the phenomenon could be further explored combining with other data.
 
@@
   scale_x_discrete() +
-  theme_classic()
+  theme_classic() +
+  theme(axis.text.x=element_text(size=6))
 
@@
 
-Other than external pattern, we are also interested how internal variable associated with bicycle theft.
+Other than external pattern, we are also interested how internal variables associated with bicycle thefts.
 @fig-diff_count_by_year shows the monthly difference in days between occurrence date and report date for each year from 2014 to 2022.
@@
 
-@tbl-rate_by_year and @tbl-rate_by_month contains recovery rate of bicycle in each year and month respectively.
+@tbl-rate_by_year and @tbl-rate_by_month contains recovery rate of bicycle thefts in each year and month respectively.
 The recovery rate is calculated as the percentage of stolen bicycle that could be successfully returned to the owner.
@@
 @tbl-rate_by_month shows that the recovery rate in winter is lower than that in summer in general.
-Consider potential seasonal pattern for bicycle theft in a year, this may also be related to the weather in Toronto that severe and freezing weather in winter constrains people's activity in both positive and negative way.
-Even recovery rate is higher in summer, it is still not a high value that largest recovery rate for each year is 17.58% and for each month is 15.70%.
+Consider potential seasonal pattern for bicycle thefts in a year, this may also be related to the weather in Toronto that severe and freezing weather in winter constrains people's activity in both positive and negative way.
+Even recovery rate is higher in summer, it is still not a high value that the largest recovery rate for each year is 17.58% and for each month is 15.70%.
+
+In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
+Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
+Although the occurrence of bicycle theft decreased recently, bicycle theft is still common in Toronto and it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
 
@@
 
-In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
-Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
-Bicycle theft is common in Toronto but it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
-
 \newpage

### 2023-11-03 outputs/paper/paper.qmd
Revise title, introduction, and data introduction

@@
 ---
-title: "Public security, law enforcement, and bicycle theft in Toronto"
+title: "Public Security Concern: Bicycle Theft in Toronto Shows a Seasonal and Geographical Pattern"
 author: "Yufei Liu"
 thanks: "Code and data are available at: https://github.com/Florence-Liu/bicycle-theft"
-date: "`r Sys.time()`"
-date-format: " MMMM YYYY"
-abstract: "Bicycle theft has been a perennial issue for city safety in Toronto.
-The analysis is conducted based on data on opendatatoronto portal including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service.
-There exists a decreasing trend and low recovery rates in overall offences in recent years even people tend to report immediately, a seasonal pattern that more offences occurred in summer, and a geographical pattern that outdoor bicycle thefts decreased.
+date: today
+date-format: long
+abstract: "Bicycle theft has been a serious issue for city safety in Toronto.
+The analysis uses data on opendatatoronto portal including occurrence date and details about Bicycle Thefts occurrences reported to Toronto Police Service.
+We find a decreasing trend and low recovery rates in overall offences in recent years even people tend to report immediately, a seasonal pattern that more offences occurred in summer, and a geographical pattern that outdoor bicycle thefts decreased.
 It implies possible ineffectiveness of tracking and security management, and further work could look at influencing factors contributing to the patterns."
@@
 Cycling is one of the most popular transportation methods and relaxing activities in Toronto and contributes to reducing traffic congestion.
-The City of Toronto is working to support cycling by building infrastructures, such as bike lanes and bike parking facilities, and create a cycling-friendly environment [@citeThecity].
-However, together with the prevalence of cycling, bicycle theft became a major concern for cyclists and a perennial problem for the city safety in Toronto.
+The City of Toronto supports cycling by building infrastructures, such as bike lanes and bike parking facilities, and create a cycling-friendly environment [@citeThecity].
+However, together with the prevalence of cycling, bicycle theft became a major concern for cyclists and the city safety in Toronto.
 
@@
 The dataset used in this analysis was obtained from Toronto's Open Data Portal and can be accessed from the package `opendatatoronto` [@citeOpendatatoronto].
-The dataset `Bicycle Thefts` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022, including detailed information about the offence such as location, bike basic information, etc [@citeData].
-The dataset has been modified several times to update the data, and in this analysis, we will use the latest `csv` version refreshed on Apr 4, 2023.
+The dataset `Bicycle Thefts` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022 [@citeData].
+We use the latest `csv` version refreshed on Apr 4, 2023 in this analysis.
 
-Due to privacy consideration, Toronto Police Service has artificially blurred the detailed location of occurrence in this data to protect parties involved and locations were offset to the nearest road intersection [@citeData].
-In this case, the accuracy of location may be impacted and become less informative.
-
-Since the occurrence of bicycle theft was self-reported, the reporter may choose to hide or inaccurately report some information such as detailed information about the bike due to privacy consideration as lost properties are not actively investigated by the Toronto Police Service and reports are only used for records purposes[@citeTorontopoliceservice].
-It is also possible that people forget the exact time of the occurrence and detailed information about the bike if things happened long time ago.
-All these factors contribute to the existence of reporting bias, and make our data biased and less accurate for some analysis.
+Since the occurrence of bicycle theft was self-reported, the reporter may choose to hide or inaccurately report some information and contributes to the reporting biases[@citeTorontopoliceservice].
+Also, the location information is blurred due to privacy consideration.
 

### 2023-11-05 outputs/paper/paper.qmd
Update introduction and reference

@@
 
-The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection, and discussion about biases and ethical concerns about the data, @sec-data_cleaning includes packages used to clean the data and cleaning process with information about variables of interest, and @sec-data_analysis includes graphs and tables representing relationships between variables, and some discussions about implications.
+The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions.
 
@@
 The dataset `Bicycle Thefts` includes all bicycle theft occurrences reported to the Toronto Police Service from 2014 to 2022 [@citeData].
+Since the occurrence of bicycle theft was self-reported, there may exist reporting bias that the reporter may choose to hide or inaccurately report some information [@citeTorontopoliceservice].
+Also, the location information is blurred due to privacy consideration.
 We use the latest `csv` version refreshed on Apr 4, 2023 in this analysis.
 
-Since the occurrence of bicycle theft was self-reported, the reporter may choose to hide or inaccurately report some information and contributes to the reporting biases[@citeTorontopoliceservice].
-Also, the location information is blurred due to privacy consideration.
-
-## Data cleaning {#sec-data_cleaning}
-
 We will use `R` to do the analysis in this report [@citeR].
-After acquiring the dataset from package `opendatatoronto`[@citeOpendatatoronto], we used package `tidyverse`[@citeTidyverse] and package `janitor`[@citeJanitor] to clean the data and select variables for further analysis.
+After acquiring the dataset from package `opendatatoronto` [@citeOpendatatoronto], we used package `tidyverse` [@citeTidyverse] and package `janitor` [@citeJanitor] to clean the data and select variables for further analysis.
 
@@
 
-* occ_year: Year offence occurred
+* `occ_year`: Year offence occurred
 
-* occ_month: Month offence occurred
+* `occ_month`: Month offence occurred
 
-* occ_date: Date offence occurred
+* `occ_date`: Date offence occurred
 
-* report_date: Date offence reported
+* `report_date`: Date offence reported
 
-* premises_type: Premises type of offence
+* `premises_type`: Premises type of offence
 
-* status: Whether the bicycle is recovered, stolen, or unknown
+* `status`: Whether the bicycle is recovered, stolen, or unknown
 
@@
 
-* occ_year: Year offence occurred
+* `occ_year`: Year offence occurred
 
-* occ_month: Month offence occurred
+* `occ_month`: Month offence occurred
 
-* difference: Difference in days between occurrence date and report date
+* `difference`: Difference in days between occurrence date and report date
 
-* location: Premises type of offence
+* `location`: Premises type of offence
 
-* bicycle_status: Whether the bicycle is recovered, stolen, or unknown
+* `bicycle_status`: Whether the bicycle is recovered, stolen, or unknown
 
-## Data analysis {#sec-data_analysis}
+## Data visualization {#sec-data_visualization}
 
@@
 #| label: fig-yearly_count_by_month
-#| fig-cap: Bar plot of monthly number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
+#| fig-cap: Monthly number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
 #| echo: false
@@
 
-data |> ggplot(aes(x=occ_year, fill=occ_month)) +
+data |> 
+  mutate(occ_month = factor(occ_month, levels = month.name)) |> 
+  ggplot(aes(x=as.factor(occ_year), fill=occ_month)) +
   geom_bar() +
   labs(x = "Occurrence Year", y = "Number of occurrence", fill = "Occurrence Month") +
-  scale_x_discrete() +
-  theme_classic()
+  theme_minimal()
 ```
@@
 #| label: fig-location_count_by_year
-#| fig-cap: Bar plot of number of occurrence of bicycle theft in difference type of location in Toronto for each year from 2014 to 2022.
+#| fig-cap: Number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
 #| echo: false

### 2023-11-13 outputs/paper/paper.qmd
Change figures and add discussion part

@@
 
-The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions.
+The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions; @sec-discussion discusses findings from data visualization and implications.
 
@@
 
-@fig-yearly_count_by_month shows monthly number of offence for each year from 2014 to 2022.
-The bar plot provides a clear demonstration for not only the overall trend of bicycle thefts through years, but also the potential seasonal pattern that could be further explored.
-It shows that the total number of offence in a year increased before 2018 and began decrease since 2018.
+```{r}
+#| label: fig-yearly_count
+#| fig-cap: Number of occurrence of bicycle theft in Toronto for each year from 2014 to 2022
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> 
+  mutate(occ_month = factor(occ_month, levels = month.name)) |> 
+  ggplot(aes(x=as.factor(occ_year))) +
+  geom_bar(color="darkgreen", fill="pink2") +
+  labs(x = "Occurrence Year", y = "Number of occurrence", fill = "Occurrence Month") +
+  theme_minimal()
+```
+
+@fig-yearly_count shows the total number of bicycle thefts in Toronto from 2014 to 2022.
+We could find that the total number of offence in a year increased before 2018 and began decrease since 2018.
 The change is relatively not significant within 1000, and the number of offence in 2022 is similar to that in 2014.
-Possible reasons include improvement of security equipment of the bike, the market behind bicycle theft influenced by pandemic, and people's usage habit of bike.
-These implications need more data to analyze and could be our further interest.
 
-@fig-yearly_count_by_month also illustrates potential seasonal pattern that there are much fewer bicycle thefts in winter, eg.
+@fig-yearly_count_by_month illustrates potential seasonal pattern that there are much fewer bicycle thefts in winter, eg.
 December, January, and February.
@@
 June, July, and August.
-This may be related to the weather in Toronto that many people will not ride bikes in winter due to extremely freezing weather and frequent snowfall.
+It also shows that the season in which bicycles were stolen differs from year to year, but overall centers in the summer seasons and the seasonal difference is significant.
 
@@
   mutate(occ_month = factor(occ_month, levels = month.name)) |> 
-  ggplot(aes(x=as.factor(occ_year), fill=occ_month)) +
-  geom_bar() +
-  labs(x = "Occurrence Year", y = "Number of occurrence", fill = "Occurrence Month") +
-  theme_minimal()
+  group_by(occ_year, occ_month) |>
+  summarise(n=n()) |>
+  ggplot(aes(x = occ_month, y = n, group = occ_year, color =  as.factor(occ_year))) +
+  geom_line(linewidth = 0.5) +
+  theme_minimal() +
+  labs(x = "Occurrence month", y = "Number of Occurrence", color = "Occurence year") +
+  theme(axis.text.x=element_text(size=6.5)) 
 ```
 
-Other than seasonal pattern, @fig-location_count_by_year also demonstrates the relationship between the location type and number of bicycle theft for each year.
+Other than seasonal pattern, @tbl-location_count demonstrates the relationship between the location type and number of bicycle theft for each year.
 It clearly shows that bicycles were mostly stolen when parked outside, and this aligns with common sense that outdoor parking is more vulnerable to theft.
 However, it is surprising that the second most common location for bicycle theft is apartment instead of house.
-Residents' transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
-For residents in house, they may rather choose vehicles that are faster such as cars to commute.
 
@@
 Factors contributing to the phenomenon could be further explored combining with other data.
-
 It is noticeable that there are many reports with location type `Other`.
@@
 ```{r}
-#| label: fig-location_count_by_year
-#| fig-cap: Number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
+#| label: tbl-location_count
+#| tbl-cap: Number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
 #| echo: false
@@
 
-data |> ggplot(aes(x=location, fill=occ_year)) +
-  geom_bar() +
-  labs(x = "Location Type", y = "Number of occurrence", fill = "Occurrence Year") +
+data |> group_by(location) |>
+  summarise(n=n()) |>
+  kable(col.names = c("Location type","Total number of occurrence"))
+```
+```{r}
+#| label: fig-location_count_by_year
+#| fig-cap: Yearly number of bicycle theft in different type of location in Toronto for each year from 2014 to 2022.
+#| echo: false
+#| warning: false
+#| message: false
+
+data |>
+  group_by(location, occ_year) |>
+  summarise(n=n()) |>
+  ggplot(aes(x=location, y=n, group=occ_year, color=as.factor(occ_year))) +
+  geom_line() +
+  labs(x = "Location Type", y = "Number of occurrence", color = "Occurrence Year") +
   scale_x_discrete() +
-  theme_classic() +
+  theme_minimal() +
   theme(axis.text.x=element_text(size=6))
@@
 @fig-diff_count_by_year shows the monthly difference in days between occurrence date and report date for each year from 2014 to 2022.
-It illustrates that people tend to report the offence within the same day of occurrence since there are more scatters near 0.
-Combined with @fig-yearly_count_by_month, there are much fewer thefts happened in December, however, the interval between occurrence and report date is much shorter in December since there are much more dots representing December near 0.
+It illustrates that people tend to report the offence within the same day of occurrence regardless of years.
+Very few people report bicycle theft after one month.
+It also consistent with @fig-yearly_count_by_month that more bicycle theft happen in summer instead of winter.
+However, the difference in report date and occurrence data does not show a seasonal pattern, namely regardless of seasons, people tend to report within the same day.
 
@@
 #| label: fig-diff_count_by_year
-#| fig-cap: Scatter plot of monthly difference in days between occurrence date and report date for bicycle thefts in Toronto for each year from 2014 to 2022.
+#| fig-cap: Monthly difference in days between occurrence date and report date for bicycle thefts in Toronto for each year from 2014 to 2022.
 #| echo: false
@@
 data |>
-  ggplot(aes(x=occ_year, y=difference,
-             color=occ_month)) +
-  geom_point(alpha = 0.5) +
-  labs(x = "Occurrence Year", y = "Difference in days between occurrence date and report date", fill = "Occurrence Month") +
-  theme_classic()
+  filter(difference <= 100) |>
+  ggplot(aes(x=difference, color=factor(occ_month, levels = month.name))) +
+  geom_freqpoly() +
+  facet_wrap(~occ_year) +
+  theme_minimal() +
+  labs(x = "Difference in days", y = "Count", color = "Occurrence month")
 ```
@@
 Consider potential seasonal pattern for bicycle thefts in a year, this may also be related to the weather in Toronto that severe and freezing weather in winter constrains people's activity in both positive and negative way.
-Even recovery rate is higher in summer, it is still not a high value that the largest recovery rate for each year is 17.58% and for each month is 15.70%.
-
-In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
-Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
-Although the occurrence of bicycle theft decreased recently, bicycle theft is still common in Toronto and it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
+Even recovery rate is higher in summer, it is still not a high value that the largest recovery rate for each year is 17.6% and for each month is 15.7%.
 
@@
 data |> group_by(occ_year) |> 
-  summarise(rate = percent(sum(bicycle_status == "RECOVERED")/nrow(data)*100)) |> 
+  summarise(rate = percent(round(sum(bicycle_status == "RECOVERED")/nrow(data)*100,3))) |> 
   arrange(rate) |> 
@@
 data |> group_by(occ_month) |> 
-  summarise(rate = percent(sum(bicycle_status == "RECOVERED")/nrow(data)*100)) |> 
+  summarise(rate = (percent(round(sum(bicycle_status == "RECOVERED")/nrow(data)*100,3)))) |> 
   arrange(rate) |> 
@@
 
+# Discussion {#sec-discussion}
+
+## Bicycle theft shows an overall decreasing trend in recent years
+
+In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
+Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
+Although the occurrence of bicycle theft decreased recently, bicycle theft is still common in Toronto and it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
+
+## Bicycle theft shows a seasonal pattern
+
+Possible reasons include improvement of security equipment of the bike, the market behind bicycle theft influenced by pandemic, and people's usage habit of bike.
+These implications need more data to analyze and could be our further interest.
+
+This may be related to the weather in Toronto that many people will not ride bikes in winter due to extremely freezing weather and frequent snowfall.
+
+## Bicycle theft shows a geographical pattern
+
+Residents' transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
+For residents in house, they may rather choose vehicles that are faster such as cars to commute.
+
 \newpage

### 2023-11-13 outputs/paper/paper.qmd
Update discussion

@@
 
-The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions; @sec-discussion discusses findings from data visualization and implications.
+The remainder of this paper is structured as follows: @sec-data discusses the data with @sec-data_introduction includes basic information about data collection and cleaning results, and @sec-data_visualization includes graphs and tables representing relationships between variables and some discussions; @sec-discussion discusses findings and implications from data visualization as well as weaknesses and future work.
 
@@
 June, July, and August.
-It also shows that the season in which bicycles were stolen differs from year to year, but overall centers in the summer seasons and the seasonal difference is significant.
+It also shows that the season in which bicycles were stolen most differs from year to year, but overall centers in the summer seasons and the seasonal difference is significant.
 
@@
   labs(x = "Occurrence month", y = "Number of Occurrence", color = "Occurence year") +
-  theme(axis.text.x=element_text(size=6.5)) 
+  theme(axis.text.x=element_text(size=6)) 
 ```
@@
 ```
+
 ```{r}
@@
 
+```{r}
+#| label: tbl-rate_by_location
+#| tbl-cap: Recovery rate of bicycle theft in Toronto from 2014 to 2022 for different location types
+#| echo: false
+#| warning: false
+#| message: false
+
+data |> group_by(location) |> 
+  summarise(rate = (percent(round(sum(bicycle_status == "RECOVERED")/nrow(data)*100,3)))) |> 
+  arrange(rate) |> 
+  kable(col.names = c("Location", "Recovery Rate"))
+
+```
+@tbl-rate_by_location shows the recovery rate of bicycle theft for different location types.
+We find that although bicycle thefts happen outside frequently as shown in @fig-location_count_by_year, these offences has the highest recovery rate.
+However, for bicycle thefts happened on transit, it is relatively hard to get the bicycle back.
+
 # Discussion {#sec-discussion}
 
-## Bicycle theft shows an overall decreasing trend in recent years
+## Bicycle theft shows an overall decreasing trend but low recovery rate in recent years
 
-In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent year with potential seasonal pattern that more offence occurred in summer than in winter, and location pattern that outdoor theft decreased but indoor theft increased in recent year with @fig-yearly_count_by_month and @fig-location_count_by_year.
-Some implications about people's reporting habit and the overall recovery rate are delivered by @fig-diff_count_by_year, @tbl-rate_by_year, and @tbl-rate_by_month.
+In general, the data shows the overall trend of bicycle theft in Toronto decreased in recent years.
+Possible reasons include improvement of security equipment of the bike and influence of pandemic on both the market behind bicycle theft and people's habit of taking bikes.
 Although the occurrence of bicycle theft decreased recently, bicycle theft is still common in Toronto and it is hard to get the stolen bikes successfully returned due to many factors such as lack of surveillance and ineffectiveness of tracking measures.
@@
 
-Possible reasons include improvement of security equipment of the bike, the market behind bicycle theft influenced by pandemic, and people's usage habit of bike.
-These implications need more data to analyze and could be our further interest.
-
+Bicycle theft in Toronto shows a seasonal pattern that more offences happened in summer seasons and fewer in winter.
+The seasonal difference is significant regardless of years.
+July used to be the season in which most bicycles were stolen across years.
 This may be related to the weather in Toronto that many people will not ride bikes in winter due to extremely freezing weather and frequent snowfall.
+However, the difference in total bicycle use is hard to track.
+The bicycles that were stolen in summer seasons, however, have a higher recovery rate comparing with 1.9% for winter seasons.
+A possible reason may also relate to weather that it is harder to track in extreme weathers.
+Also, it is noticeable that the total bicycle theft as well as possible total bicycle use in winter are both lower than that in summer, thus possibly making the rate to be low.
 
@@
 
+Bicycle theft in Toronto also shows a geographical pattern that bicycles parked outside are more likely to be stolen.
+For inside locations, apartment is the most common location that the bicycles were stolen.
 Residents' transportation habits may be a possible reason that since apartments are usually located in the town center and close to public transportation stations, bicycles is a good choice for short trips.
 For residents in house, they may rather choose vehicles that are faster such as cars to commute.
+For other types of locations, there may exist plausible difference in total bicycle use for different locations, which is also hard to track.
+
+## Weaknesses and future work
+
+One weakness of the paper is about data.
+Since bicycle thefts were self-reported, there exist reporting biases that the reporter may choose to hide or accidentally forget some information.
+Also, data about location was blurred due to privacy consideration.
+Additionally, the data is not up-to-date that it does not include data for 2023.
+Another point is that the total bicycle use in Toronto is missed so we could not compare the total bicycle use.
+This is critical for determining the geographical pattern of bicyle theft as mentioned above.
+It is plausible that people tend to use bicycle less in certain locations.
+
+Further work could focus on using more data and include look at other influencing factors contributing to bicycle theft such as bicycle types and brands.
+However, it could be helpful if the data is more credible than self-reporting.
+We may also look at possible factors behind the seasonal pattern and geographical pattern other than weather.
 
